Commutative character sheaves and geometric types for supercuspidal representations

### 2017-04-06 Geometric types.tex
Finished the proof of Theorem 3.4.

\begin{proof}
Notation as above.
Consider the following diagram.
%\[
%\begin{tikzcd}
%{} & G\times G \arrow{d}{c} &&&\\
%1 \arrow{r} & H \arrow{r}{j} & G \arrow{r}{q} & G\ab \arrow{r} & 1
%\end{tikzcd}
%\]
%gives
\[
\begin{tikzcd}
\CS(G\der) & \arrow{l}[swap]{j^*} \CS(G) &\arrow{l}[swap]{q^*} 
\CS(G\ab) \arrow{dl}{\text{equiv}}[swap]{q\ab^*} \\
& \CSab(G) \arrow{u}{\text{forget}}  & 
\end{tikzcd}
\]

@@

Conversely, suppose $(\cs{L},\mu)\in \CS(G)$ and $(\cs{L},\mu) \iso q^*(\cs{L}\ab,\mu\ab)$ in $\CS(G)$.
Then $j^*(\cs{L},\mu) \iso j^*q^*(\cs{L}\ab,\mu\ab)$ in $\CS(G\der)$. 
Since $q\circ j = 1$, it follows that $\cs{L}\vert_{G\der} \iso (\EE)_{G\der}$ in $\CS(G\der)$.
\end{proof}

@@

We may now define commutative character sheaves on $G$.

Every $\beta : \cs{L}\vert_{G\der} \to (\EE)_{G\der}$ determines an isomorphism $\gamma : m^*\cs{L} \to \theta^*m^*\cs{L}$ as follows.
Let $i : G \to G$ be inversion and $c : G\times G\to G\der$ be the commutator map, defined by $c(x,y)= xyx^{-1}y^{-1}$.
Both are smooth morphisms of $\Fq$-schemes.
Set $m' = i \circ m \circ \theta$; we have $j\circ c = m \circ (m \times m')$.
Then $\beta : \cs{L}\vert_{G\der} \to (\EE)_{G\der}$ determines the isomorphism $\gamma' : m^*\cs{L} \otimes \theta^* m^* i^*\cs{L} \to (\EE)_{G\times G}$ by the diagram of isomorphisms below.
\begin{equation}
\begin{tikzcd}
\arrow[equal]{d} c^* (\cs{L}\vert_{G\der}) \arrow{r}{c^*(\beta)} 
	&  c^*((\EE)_{G\der}) \arrow[equal]{d} \\
\arrow{d}[swap]{j\circ c = m \circ (m \times m')} c^* j^* \cs{L} 
	&   (\EE)_{G\times G} \\
(m \times m')^* m^* \cs{L} \arrow{d}[swap]{(m \times m')^*(\mu)} 
	&  m^*\cs{L} \otimes \theta^* m^* i^* \cs{L}  \arrow[dashed]{u}[swap]{\gamma'} \\
(m \times m')^* (\cs{L} \boxtimes \cs{L}) \arrow[equal]{r} 
	& m^*\cs{L} \otimes (m')^*\cs{L} \arrow{u}[swap]{m' = i\circ m\circ \theta} 
\end{tikzcd}
\end{equation}
In the diagram above, the arrows labeled with equations come from canonical isomorphisms of functors on Weil sheaves derived from the equations; so, for example, the middle left isomorphism comes from $(m\times m')^* m^* \iso c^* j^*$ since $j\circ c = m \circ (m \times m')$.
Using the monoidal structure of the category of Weil local systems on $G\times G$, the isomorphism $\gamma' : m^*\cs{L} \otimes \theta^* m^* i^*\cs{L} \to (\EE)_{G\times G}$ defines an isomorphism
\[
m^*\cs{L} \to (\theta^* m^* i^*\cs{L})^\vee.
\]
Applying the canonical isomorphisms $(\theta^* m^* i^*\cs{L})^\vee \iso \theta^* m^* i^* (\cs{L}^\vee)$ and $i^*(\cs{L}^\vee) \cong \cs{L}$, this map provides the promised isomorphism
\[
\begin{tikzcd}
\gamma : m^*\cs{L} \arrow{r} & \theta^* m^* \cs{L}.
\end{tikzcd}
\]

@@

\begin{proof}
We again write $H$ for $G\der$.
By Definition~\ref{def:CCScom}, $\CCS(G\ab) \to \CS(G\ab)$ is a full subcategory.
Since the comparison functor $\CS(G\ab) \to \CS_{H}(G)$ is an equivalence, it determines a full subcategory $\CCS_{H}(G)$ which is equivalent to $\CCS(G\ab)$, as pictured below.
The proof of Theorem~\ref{thm:Gab} now reduces to the following claim: the essential image of $\CCS_H(G)$ under the equivalence $\CS_{H}(G) \to \CSab(G)$ is $\CCS(G)$.
\[
\begin{tikzcd}[row sep=30, column sep=40]
\CSab(G) & \CS(G\ab) \arrow{l}{\text{equiv}}[swap]{q\ab^*} \\
\CCS(G) \arrow{u}{\text{full sub}} & \CCS(G\ab) \arrow{u}[swap]{\text{full sub}} \arrow{l}[swap]{\text{equiv}}
\end{tikzcd}
%\begin{tikzcd}[row sep=30, column sep=40]
%\CS(H) 
%& \arrow{l}[swap]{j^*} \CS(G) 
%&\arrow{l}[swap]{q^*} 
%\CS(G\ab) \arrow{dl}{\text{equiv}}[swap]{q_{H}^*} \\
%\CS'(G)  \arrow{r}{\beta \mapsto \nu}[swap]{\text{equiv}} 
%&  \CS_{H}(G) \arrow{u}{\text{forget}}  
%&  %\arrow[bend left=35pt, dashed]{dll} 
%\CCS(G\ab) \arrow{u}[swap]{\text{full sub}} \arrow{dl}{\text{equiv}} \\
%\CCS(G) \arrow{u}{\text{full sub}} \arrow{r}{\beta \mapsto \nu}[swap]{\text{equiv}}
%& \arrow{u}{\text{full sub}}  \CCS_{H}(G) 
%& 
%\end{tikzcd}
\]

Let $(\cs{L},\mu,\nu)$ be the image of $(\cs{L},\mu,\beta)\in \CSab(G)$ and of $(\cs{L}\ab,\mu\ab)$ under $q_H^*$;
we must show that $(\cs{L},\mu,\beta)\in \CCS(G)$ if and only if $(\cs{L}\ab,\mu\ab)\in \CCS(G\ab)$.
Let $\xi : m\ab^*\cs{L}\ab \to \theta^* m\ab^*\cs{L}\ab$ be the isomorphism attached to $(\cs{L}\ab,\mu\ab)\in \CS(G\ab)$ as in Section~\ref{sec:comcom}.
Let $\gamma : m^*\cs{L} \to \theta^* m^*\cs{L}$ be the isomorphism attached to $\beta : \cs{L}\vert_{H} \to (\EE)_{H}$ as in Section~\ref{ssec:noncomdef}.
Then the diagram in Definition~\ref{def:CCS} is precisely the result of applying the functor $(q\times q)^*$ to the diagram in Definition~\ref{def:CCScom}, as pictured below; 
in particular $\gamma = (q\times q)^* \xi$.
  \[
  \begin{tikzcd}[row sep=10, column sep = 20]
   m_0^*\cs{L}_0 \arrow{dd}[swap]{\xi} \arrow{r}{\mu_0} 
   & \cs{L}_0\boxtimes\cs{L}_0 \arrow{dd}{\vartheta}
&&&   m^*\cs{L} \arrow{dd}[swap]{\gamma} \arrow{r}{\mu} 
 & \cs{L}\boxtimes\cs{L} \arrow{dd}{\vartheta}\\
 && \arrow{r}{(q\times q)^*}  &\ & & &\\ 
   \theta^*(m_0^*\cs{L}_0) \arrow{r}{{\theta}^*\mu_0} 
 &  \theta^*(\cs{L}_0\boxtimes\cs{L}_0)
&&&   \theta^*(m^*\cs{L}) \arrow{r}{{\theta}^*\mu} 
&  \theta^*(\cs{L}\boxtimes\cs{L})
  \end{tikzcd}
\]
Using Lemma~\ref{lem:HH}, we may interpret the diagram on the right as a diagram in $\Loc_{H\times H}(G\times G)$.
By Lemma~\ref{lem:torsor}, this corresponds to a diagram in $\Loc(G\ab\times G\ab)$, necessarily the diagram on the left, above, and also that the diagram in Definition~\ref{def:CCS} commutes if and only if the diagram in Definition~\ref{def:CCScom} commutes. 
In other words, $(\cs{L},\mu,\beta)\in \CCS(G)$
if and only if $(\cs{L}\ab,\mu\ab)\in \CCS(G)$.

\end{proof}

>>>>

\begin{proof}
Notation as in the proof of Proposition~\ref{prop:CSGabab}.
Consider the following diagram.
%\[
%\begin{tikzcd}
%{} & G\times G \arrow{d}{c} &&&\\
%1 \arrow{r} & H \arrow{r}{j} & G \arrow{r}{q} & G\ab \arrow{r} & 1
%\end{tikzcd}
%\]
%gives
\[
\begin{tikzcd}
\CS(G\der) & \arrow{l}[swap]{j^*} \CS(G) &\arrow{l}[swap]{q^*} 
\CS(G\ab) \arrow{dl}{\text{equiv}}[swap]{q\ab^*} \\
& \CSab(G) \arrow{u}{\text{forget}}  & 
\end{tikzcd}
\]

@@

Conversely, suppose $(\cs{L},\mu)\in \CS(G)$ and $(\cs{L},\mu) \iso q^*(\cs{L}\ab,\mu\ab)$ in $\CS(G)$.
Then $j^*(\cs{L},\mu) \iso j^*q^*(\cs{L}\ab,\mu\ab)$ in $\CS(G\der)$. 
Since $q\circ j = 1_{G\ab}$, it follows that $\cs{L}\vert_{G\der} \iso (\EE)_{G\der}$ in $\CS(G\der)$.
\end{proof}

@@

We may now define commutative character sheaves on $G$.
%
Every $\beta : \cs{L}\vert_{G\der} \to (\EE)_{G\der}$ determines an isomorphism $\gamma : m^*\cs{L} \to \theta^*m^*\cs{L}$ as follows.
Let $i : G \to G$ be inversion and $c : G\times G\to G\der$ be the commutator map, defined by $c(x,y)= xyx^{-1}y^{-1}$.
Both are smooth morphisms of $\Fq$-schemes.
Set $m' = i \circ m \circ \theta$; we have $j\circ c = m \circ (m \times m')$.
Then $\beta : \cs{L}\vert_{G\der} \to (\EE)_{G\der}$ determines the isomorphism $\gamma' : m^*\cs{L} \otimes \theta^* m^* i^*\cs{L} \to (\EE)_{G\times G}$ by the diagram of isomorphisms below.
\begin{equation}
\begin{tikzcd}
\arrow[equal]{d} c^* (\cs{L}\vert_{G\der}) \arrow{r}{c^*(\beta)} 
	&  c^*((\EE)_{G\der}) \arrow[equal]{d} \\
\arrow{d}[swap]{j\circ c = m \circ (m \times m')} c^* j^* \cs{L} 
	&   (\EE)_{G\times G} \\
(m \times m')^* m^* \cs{L} \arrow{d}[swap]{(m \times m')^*(\mu)} 
	&  m^*\cs{L} \otimes \theta^* m^* i^* \cs{L}  \arrow[dashed]{u}[swap]{\gamma'} \\
(m \times m')^* (\cs{L} \boxtimes \cs{L}) \arrow[equal]{r} 
	& m^*\cs{L} \otimes (m')^*\cs{L} \arrow{u}[swap]{m' = i\circ m\circ \theta} 
\end{tikzcd}
\end{equation}
In the diagram above, the arrows labeled with equations come from canonical isomorphisms of functors on Weil sheaves derived from the equations; so, for example, the middle left isomorphism comes from $(m\times m')^* m^* \iso c^* j^*$ since $j\circ c = m \circ (m \times m')$.
Using the monoidal structure of the category of Weil local systems on $G\times G$, the isomorphism $\gamma' : m^*\cs{L} \otimes \theta^* m^* i^*\cs{L} \to (\EE)_{G\times G}$ defines an isomorphism
\[
m^*\cs{L} \to (\theta^* m^* i^*\cs{L})^\vee.
\]
Applying the canonical isomorphisms $(\theta^* m^* i^*\cs{L})^\vee \iso \theta^* m^* i^* (\cs{L}^\vee)$ and $i^*(\cs{L}^\vee) \cong \cs{L}$, this map provides the promised isomorphism
\[
\begin{tikzcd}
\gamma : m^*\cs{L} \arrow{r} & \theta^* m^* \cs{L}.
\end{tikzcd}
\]

@@

\begin{proof}
%We again write $H$ for $G\der$.
By Definition~\ref{def:CCScom}, $\CCS(G\ab) \to \CS(G\ab)$ is a full subcategory.
Since the comparison functor $\CS(G\ab) \to \CS_{G\der}(G)$ is an equivalence, it determines a full subcategory $\CCS_{G\der}(G)$ which is equivalent to $\CCS(G\ab)$, as pictured below.
The proof of Theorem~\ref{thm:Gab} now reduces to the following claim: the essential image of $\CCS_{G\der}(G)$ under the equivalence $\CS_{G\der}(G) \to \CSab(G)$ is $\CCS(G)$.
\[
\begin{tikzcd}[row sep=30, column sep=40]
\CSab(G) & \CS(G\ab) \arrow{l}{\text{equiv}}[swap]{q\ab^*} \\
\CCS(G) \arrow{u}{\text{full sub}} & \CCS(G\ab) \arrow{u}[swap]{\text{full sub}} \arrow{l}[swap]{\text{equiv}}
\end{tikzcd}
%\begin{tikzcd}[row sep=30, column sep=40]
%\CS(H) 
%& \arrow{l}[swap]{j^*} \CS(G) 
%&\arrow{l}[swap]{q^*} 
%\CS(G\ab) \arrow{dl}{\text{equiv}}[swap]{q_{H}^*} \\
%\CS'(G)  \arrow{r}{\beta \mapsto \nu}[swap]{\text{equiv}} 
%&  \CS_{H}(G) \arrow{u}{\text{forget}}  
%&  %\arrow[bend left=35pt, dashed]{dll} 
%\CCS(G\ab) \arrow{u}[swap]{\text{full sub}} \arrow{dl}{\text{equiv}} \\
%\CCS(G) \arrow{u}{\text{full sub}} \arrow{r}{\beta \mapsto \nu}[swap]{\text{equiv}}
%& \arrow{u}{\text{full sub}}  \CCS_{H}(G) 
%& 
%\end{tikzcd}
\]

Let $(\cs{L},\mu,\nu)$ be the image of $(\cs{L},\mu,\beta)\in \CSab(G)$ and of $(\cs{L}\ab,\mu\ab)$ under $q\ab^*$;
we must show that $(\cs{L},\mu,\beta)\in \CCS(G)$ if and only if $(\cs{L}\ab,\mu\ab)\in \CCS(G\ab)$.
Let $\xi : m\ab^*\cs{L}\ab \to \theta^* m\ab^*\cs{L}\ab$ be the isomorphism attached to $(\cs{L}\ab,\mu\ab)\in \CS(G\ab)$ as in Section~\ref{sec:comcom}.
Let $\gamma : m^*\cs{L} \to \theta^* m^*\cs{L}$ be the isomorphism attached to $\beta : \cs{L}\vert_{H} \to (\EE)_{H}$ as in Section~\ref{ssec:noncomdef}.
Recall the equivalence $L_2 : \Loc(G^2\ab) \to  \Loc_{G^2\der}(G^2)$ from Proposition~\ref{prop:CSGabab}.
The diagram in Definition~\ref{def:CCS} is precisely the result of applying the functor $L_2$ to the diagram in Definition~\ref{def:CCScom}, as pictured below; 
in particular $\gamma = (q^2)^* \xi$.
  \[
  \begin{tikzcd}[row sep=10, column sep = 20]
   m_0^*\cs{L}_0 \arrow{dd}[swap]{\xi} \arrow{r}{\mu_0} 
   & \cs{L}_0\boxtimes\cs{L}_0 \arrow{dd}{\vartheta}
&&&   m^*\cs{L} \arrow{dd}[swap]{\gamma} \arrow{r}{\mu} 
 & \cs{L}\boxtimes\cs{L} \arrow{dd}{\vartheta}\\
 && \arrow{r}{(q^2)^*}  &\ & & &\\ 
   \theta^*(m_0^*\cs{L}_0) \arrow{r}{{\theta}^*\mu_0} 
 &  \theta^*(\cs{L}_0\boxtimes\cs{L}_0)
&&&   \theta^*(m^*\cs{L}) \arrow{r}{{\theta}^*\mu} 
&  \theta^*(\cs{L}\boxtimes\cs{L})
  \end{tikzcd}
\]
The diagram on the right is a diagram in $\Loc_{G^2\der}(G^2)$.
Using the equivalence $L_2$, this corresponds to a diagram in $\Loc(G^2\ab)$, necessarily the diagram on the left, above.
It follows that the diagram in Definition~\ref{def:CCS} commutes if and only if the diagram in Definition~\ref{def:CCScom} commutes. 
In other words, $(\cs{L},\mu,\beta)\in \CCS(G)$
if and only if $(\cs{L}\ab,\mu\ab)\in \CCS(G)$.
\end{proof}
